Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

    \figcapup
    \caption{Suppose $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
    \label{fig:guard}
    \figcapdown
\end{figure}

\section{Geometric Building Bricks} \label{sec:bricks}

This section will first define some notions frequently used in our presentation and then introduce several computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.

\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$. Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$. We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

    \item $\gbot(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
}
See Figure~\ref{fig:guard} for

\begin{figure*}
    \begin{tabular}{ccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} \\
%         &
%         \hspace{3mm}
%         \includegraphics[height=30mm]{./artwork/prob-d} \\
        (a) Problems $\EuScript{A}$ and $\EuScript{D}$ &
        \hspace{3mm}
        (b) Problem $\EuScript{B}$ &
        \hspace{3mm}
        (c) Problem $\EuScript{C}$
%         &
%         \hspace{3mm}
%         (d) Problem $\EuScript{D}$
    \end{tabular}

@@

@@

\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.

@@

\section{Algorithm for Problem $\bm{\EuScript{A}}$} \label{app:prob-a}

\section{Algorithm for Problem $\bm{\EuScript{B}}$} \label{app:prob-b}

\section{Algorithm for Problem $\bm{\EuScript{C}}$} \label{app:prob-c}

\section{Algorithm for Problem $\bm{\EuScript{D}}$} \label{app:prob-d}

>>>>

    \figcapup
    \caption{Define 3-tuple $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
    \label{fig:guard}
    \figcapdown
\end{figure}

\section{Preliminaries in Geometry} \label{sec:bricks}

This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.

\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$. Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?} We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$. We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

    \item $\gbot(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
}
See Figure~\ref{fig:guard} for an illustration.

\begin{figure*}
    \begin{tabular}{cccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=27mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-d} \\[2mm]
        (a) Problems $\EuScript{A}$ &
        \hspace{3mm}
        (b) Problem $\EuScript{B}$ &
        \hspace{3mm}
        (c) Problem $\EuScript{C}$
        &
        \hspace{3mm}
        (d) Problem $\EuScript{D}$
    \end{tabular}

@@

\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space. In the {\em detection version} of Problem $\EuScript{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$. Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.

In the {\em reporting version} of the problem, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$, provided that at least one such $r$ exists. For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.

\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space. The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$. If $h$ does not intersect with any segment in $V$, report nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.

\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.

@@

@@

\section{Building Brick Geometry Algorithms} \label{app:bricks}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{A}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{B}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{C}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{D}}$}
